Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-29 overview.tex
Added a table of variables and constants used in the paper.

We begin this section by first writing down the complete set of
physical equations that can be solved by \enzo, and then briefly
describe the numerical algorithms that we use to solve these
equations.  This section is intended to be an overview of the \enzo\
code's capabilities: thus, we gather all of the equations solved into
a single place and provide a brief and high-level introduction to the
numerics of the code.  Detailed descriptions of the individual
components are then provided in
Sections~\ref{sec.amr}-\ref{sec.num.analysis}.

@@

%\begin{eqnarray} 
%\frac{\partial \rho}{\partial t} 
%          + \vecv \cdot \nabla \rho
%    & = & - \rho \nabla \cdot \vecv,
%        \label{eq:mass} \\
%%
%\frac{\partial \vecv }{ \partial t} 
%          + ( \vecv \cdot \nabla ) \vecv 
%    & = & - \frac{1}{\rho} \nabla p
%          - \nabla \phi,
%        \label{eq:momentum} \\
%%
% \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
%    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
%          - \vecv \cdot \nabla \phi
%%\nonumber          \\  & &     % uncomment for 2-column
%     + \Gamma - \Lambda
%        \label{eq:total_energy}
%\end{eqnarray}
%
In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
total fluid energy, comoving baryonic density, peculiar velocity, and
comoving magnetic field strength, respectively.  $\myvec{I}$ is the
identity matrix, and $a$ is the cosmological expansion parameter
(which is discussed in more detail below).  The first equation
represents conservation of mass, the second, conservation of momentum,
and the third, conservation of total fluid (kinetic plus thermal)
energy.  They are respectively, the first, second and third moments of
the Boltzmann equation.  The fourth equation is the magnetic induction
equation.  Terms representing radiative cooling ($\Lambda$) and
heating ($\Gamma$) enter on the right-hand side of the energy equation
(\ref{eq:total_energy}), as well as the flux due to thermal heat
conduction ($\fcond$).

The total fluid energy $E$ is given by:
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def}.
\end{equation}
where $e$ is the comoving thermal energy density, and the total
isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a},
\end{equation}
where $p$ is the gas pressure.  The equations are closed with the
equation of state and Poisson's equation (we work in the Newtonian
limit):
%
\begin{eqnarray}
%
  e &=& p / [(\gamma - 1) \rho ],
  \label{eq:eq_of_state} \\
%
  \nabla^2 \phi &=& 4 \pi G \rho_{\rm total}
%                   + 3 p_{total}/c^2) - \Lambda. 
  \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of
specific heats $\gamma$.  Equation (\ref{eq:potential}) includes
contributions to the gravitational potential $\phi$ from the total
mass density $\rho_{\rm total}$, which is defined as $\rho_{\rm gas} +
\rho_{\rm dm} + \rho_{\rm stars} - \rho_{\rm average}$.  Although we
write the equations including both MHD and comoving coordinates, the
code is frequently used both in the pure hydrodynamic limit ($\vecB =
0$) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).

>>>>

We begin this section by first writing down the complete set of
physical equations that can be solved by \enzo, and then briefly
describe the numerical algorithms that we use to solve these
equations.  This section is intended to be an overview of the
\enzo\ code's capabilities: thus, we gather all of the equations
solved into a single place and provide a brief and high-level
introduction to the numerics of the code.  Detailed descriptions of
the individual components are then provided in
Sections~\ref{sec.amr}-\ref{sec.num.analysis}. In
Table~\ref{tab:variables} we provide a convenient summary of all
variables and physical constants used throughout this manuscript.

\mqk{Move this table to the Appendix?}
\input{table_of_variables_and_constants}

@@

%\begin{eqnarray} 
%\frac{\partial \rho}{\partial t} 
%          + \vecv \cdot \nabla \rho
%    & = & - \rho \nabla \cdot \vecv,
%        \label{eq:mass} \\
%%
%\frac{\partial \vecv }{ \partial t} 
%          + ( \vecv \cdot \nabla ) \vecv 
%    & = & - \frac{1}{\rho} \nabla p
%          - \nabla \phi,
%        \label{eq:momentum} \\
%%
% \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
%    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
%          - \vecv \cdot \nabla \phi
%%\nonumber          \\  & &     % uncomment for 2-column
%     + \Gamma - \Lambda
%        \label{eq:total_energy}
%\end{eqnarray}
%
In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
total fluid energy density, comoving baryonic density, peculiar velocity, and
comoving magnetic field strength, respectively.  $\myvec{I}$ is the
identity matrix, and $a$ is the cosmological expansion parameter
(which is discussed in more detail below).  The first equation
represents conservation of mass, the second, conservation of momentum,
and the third, conservation of total fluid (kinetic plus thermal)
energy.  They are respectively, the first, second and third moments of
the Boltzmann equation.  The fourth equation is the magnetic induction
equation.  Terms representing radiative cooling ($\Lambda$) and
heating ($\Gamma$) enter on the right-hand side of the energy equation
(\ref{eq:total_energy}), as well as the flux due to thermal heat
conduction ($\fcond$).

The total fluid energy density $E$ is given by:
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def}.
\end{equation}
where $e$ is the comoving thermal energy density, and the total
isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a},
\end{equation}
where $p$ is the gas pressure.  The equations are closed with the
equation of state and Poisson's equation (we work in the Newtonian
limit):
%
\begin{eqnarray}
%
  e &=& p / [(\gamma - 1) \rho ],
  \label{eq:eq_of_state} \\
%
  \nabla^2 \phi &=& 4 \pi G \rho_{\rm total}
%                   + 3 p_{total}/c^2) - \Lambda. 
  \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of
specific heats $\gamma$.  Equation (\ref{eq:potential}) includes
contributions to the gravitational potential $\phi$ from the total
mass density $\rho_{\rm total}$, which is defined as $\rho_{\rm gas} +
\rho_{\rm dm} + \rho_{\rm stars} - \rho_{\rm average}$.  Although we
write the equations including both MHD and comoving coordinates, the
code is frequently used both in the pure hydrodynamic limit ($\vecB =
0$) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).

### 2013-05-29 tests-onezonecollapse.tex
Added a table of variables and constants used in the paper.

with G being the gravitational constant. \mqk{Instead of explaining what $G$ is here, would it make sense to have a table at the beginning of the paper listing all the natural constants and variable symbols that are used throughout the paper?} The collapse timescale is
altered from the dynamical time by a factor $1/\sqrt{1-f}$ in Equation
\ref{eqn.tcol}, which is an approximation of the ratio of the gas
pressure to the force of gravity.  The value of $f$ depends on the
effective adiabatic index, $\gamma_{\rm ef} \equiv (\partial \ln p
/ \partial \ln \rho)$, which we linearly extrapolate
from derivative values at the two previous timesteps.  For the value
of $f$ in this test problem, we use the piecewise function of
\citet{2005ApJ...626..627O}, given by

>>>>

 The collapse timescale is altered from the dynamical time by a factor
$1/\sqrt{1-f}$ in Equation
\ref{eqn.tcol}, which is an approximation of the ratio of the gas
pressure to the force of gravity.  The value of $f$ depends on the
effective adiabatic index, $\gamma_{\rm ef} \equiv (\partial \ln p
/ \partial \ln \rho)$, which we linearly extrapolate
from derivative values at the two previous timesteps.  For the value
of $f$ in this test problem, we use the piecewise function of
\citet{2005ApJ...626..627O}, given by
